---
title: "WA. Introduction to Networks"
author: "Dr. Uma Ravat"
subtitle: '<br><br>Introduction to Social Network Analysis in R '
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


# 1. Introduction to Networks

## -------~~ 1.Set up --------

```{r load_packages}
# (remove # below and run this if they are not already installed )
# once installed put # back to avoid warnings/errors
#install.packages("igraph")
#install.packages("igraphdata")

library(igraphdata)
library(igraph)

```

`r bfcolor("Question:", "blue")` What is the command to load a dataset? Load a library

`r bfcolor("Solution:", "blue")` 


Some setup so plots are bigger with no margins

``` {r}

#default margins
def_marg <- c(5.1, 4.1, 4.1, 2.1)

#no margins
no_marg <- c(0, 0, 0, 0)

```


# 3. Your Turn: Explore the basic characteristics of the karate network


```{r}
# Load the karate network and read the help page
#Write your code here

```

- `r bfcolor("Question:", "blue")` Briefly describe what the karate network is about. 


`r bfcolor("Solution:", "blue")` 





```{r}
# Explore the general characteristics and attributes of the karate network 
# Write your code here
```

Use the above output of characteristics and attributes of the karate network to answer the questions below. Add code chunks as necessary to run your code and answer the questions below

- `r bfcolor("Question:", "blue")` How many nodes does the network have?

`r bfcolor("Solution:", "blue")` 


- `r bfcolor("Question:", "blue")` Is the network directed or undirected?


`r bfcolor("Solution:", "blue")` 


- `r bfcolor("Question:", "blue")` What vertex and edge attributes does the network have?

`r bfcolor("Solution:", "blue")` 

+ `r bfcolor("Question:", "blue")` Does the graph have multiple edges per pair of nodes or any loops?

`r bfcolor("Solution:", "blue")` 



# 6. Your turn! Take a closer look at Karate network!

**To answer each question in code chunks below**,

1. Replace every [xxx] with your code
2. Run the code chunk individually by clicking the "green triangle"
3. To see your solution in the knitted document, make sure to replace `eval=FALSE` with `eval=TRUE`

`r bfcolor("Question:", "blue")` How many factions does the karate network have? 

`r bfcolor("Solution:", "blue")` 


`r bfcolor("Question:", "blue")` How many nodes belong to each faction?

`r bfcolor("Solution:", "blue")` 

```{r, eval=FALSE}
table(V(karate)$[xxx])
```

`r bfcolor("Question:", "blue")` Which nodes nominated Mr. Hi?

`r bfcolor("Solution:", "blue")` 

```{r, eval=FALSE}

neighbors(karate, "Mr Hi", mode = [xxx])

```

`r bfcolor("Question:", "blue")` Which nodes nominated John A?

`r bfcolor("Solution:", "blue")` 

```{r, , eval=FALSE}
[xxx](karate, "John A", mode = "in")
```

`r bfcolor("Question:", "blue")` How many nodes nominated both Mr Hi and John A?


`r bfcolor("Solution:", "blue")` 

```{r, , eval=FALSE}

mh <- neighbors([xxx], "Mr Hi", mode = [xxx])
ja <- neighbors([xxx], "John A", mode = [xxx])

sum(mh %in% ja)

```

`r bfcolor("Question:", "blue")` Is there an edge between Mr Hi and John A?

`r bfcolor("Solution:", "blue")` 

```{r, eval=FALSE}
E(karate)["Mr Hi" [xxx] "John A"]
```

