---
title: "3 Activity: Descriptive Network Statistics (Metrics) "
author: "Dr. Uma Ravat"
subtitle: '<br><br>Introduction to Social Network Analysis in R '
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


# 1. Setting up the network 

Load the `UKFaculty` Network

```{r setup_network}
# (remove # below and run this if they are not already installed )
# once installed put # back to avoid warnings/errors
#install.packages("igraph")
#install.packages("igraphdata")

library(igraphdata)
library(igraph)

## Some setup so plots are bigger with no margins
#default margins
def_marg <- c(5.1, 4.1, 4.1, 2.1)

#no margins
no_marg <- c(0, 0, 0, 0)

## load the data
data(UKfaculty)
```



### Activity 1:  Node or Network Metrics? 

List  each metric we looked at below. Discuss with your peers, whether it is a node level(individual) network metric or a whole network metric? 


### Activity 2: Descriptive statistics for the karate network

Obtain each of the descriptive statistics for the `karate` network and interpret what these metrics say about the individuals as well as the whole network. 

_Be sure to load the network first_

_Add code chunks as necessary to run your code and answer the following_

`r bfcolor("Question:", "blue")`

```{r}
# Metric 1
#Your code here
```

Your Interpretation of Metric 1 here: 


```{r}
# Metric 2
#Your code here
```

Your Interpretation of Metric 2 here: 


## Activity 3: Use the help for each function above and use the metrics with other agruments to explore the `karate` network and calculate and interpret other aspects of the `karate` network. 