---
title: "2 WA. Network Visualization "
author: "Dr. Uma Ravat"
subtitle: '<br><br>Introduction to Social Network Analysis in R '
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


# 1. Setting up the network 

Load the `UKFaculty` Network

```{r setup_network}
# (remove # below and run this if they are not already installed )
# once installed put # back to avoid warnings/errors
#install.packages("igraph")
#install.packages("igraphdata")

library(igraphdata)
library(igraph)

## Some setup so plots are bigger with no margins
#default margins
def_marg <- c(5.1, 4.1, 4.1, 2.1)

#no margins
no_marg <- c(0, 0, 0, 0)

## load the data
data(UKfaculty)
```


# 2. Network visualization

- via algorithms

---

### Basic plot

```{r}
#basic plot
plot(UKfaculty)
```

---

### Plotting Parameters : Nodes 

vertex.[parameter] (e.g., vertex.color =)

```{r, echo=FALSE}
dfn <- data.frame(node_parameter = c("color", 
                               "shape",
                               "size", 
                               "label", 
                               "label family",
                               "label font",
                               "label cex", 
                               "label distance", 
                               ""), 
                  purpose_options =c("node color", "circle, square, rectangle, sphere, pie, raster, etc.", 
                                    "default = 15", "vector with labels", "font family of the label: Helvetica, Times", "1= plain, 2= bold, 3= italic, 4 = bold italic", "font size", "distance between node and label", ""))
dfn %>%
   knitr::kable(format = "html")
```



---

###  Plotting Parameters: Edges

edge.[parameter] (e.g., edge.color =)

```{r, echo=FALSE}
dfe <- data.frame(edge_parameter = c("color", 
                               "lty",
                               "width", 
                               "label",
                               "label family",
                               "label font",
                               "label cex", 
                               "curvature", 
                               "arrow size"),
                  purpose_options =c("edge color", "type of line: 0='blank', 1= 'solid', 2= 'dashed', 3= 'dotted', 4 = 'dotdash', 5= 'longdash', 6= 'twodash'","default = 1", "vector with labels", "font family of the label: Helvetica, Times", "1= plain, 2= bold, 3= italic, 4 = bold italic", "font size", "edge curvature - range [0,1]", "default = 1"))

dfe %>%
   knitr::kable(format = "html")

```

---

### Aside about colors 

R has 657 built in color names!

To see a list of names:
```{r}
colors()
```

### Improving the plot

+ Reduce the crossing of edges 
+ Keep the edge length as constant as possible

---



#### **Option 1:** Setting visual properties outside the plot function

We add attributes to the network

**Step 1** Set vertex and/or edge attributes

```{r}

# vertex properties 
V(UKfaculty)$color <- "#88398A"

# vertex label properties
V(UKfaculty)$label.color <- "white"
V(UKfaculty)$label.cex <- 0.5

# edge properties
E(UKfaculty)$color <- "gray"

```

**Step 2** Plot network

```{r}

plot(UKfaculty)

```

---

#### Option 2 Setting visual properties inside the plot function

We use the function arguments

```{r}

plot(UKfaculty, 
     # === vertex properties 
     vertex.color = "#88398A",
     vertex.frame.color = "#88398A",
     vertex.size = 9,
     # === vertex label properties
     vertex.label.cex = 0.5,
     vertex.label.color = "white",
     vertex.label.family = "Helvetica",
     # === edge properties
     edge.color = "gray",
     edge.width = 0.5,
     edge.arrow.size = 0.2)
```


---

#### Which option to use? 

1. Option 1(outside the plot function) : Use for assigning different visual properties to subsets of edges or nodes. 
2. Option 2(inside the plot function): Use for assigning global visual properties to all edges or nodes.

Most often, combine the two


### Coloring nodes by group to which they belong

```{r}
table(V(UKfaculty)$Group)

mycolors <- c("black", 
              "#D3D3D3", 
              "#88398A", 
              "white", 
              "#562457")
mylabcols <- c("white", "black", "white", "black", "white")

V(UKfaculty)$color <- mycolors[V(UKfaculty)$Group]
V(UKfaculty)$frame.color <- mycolors[V(UKfaculty)$Group]
V(UKfaculty)$label.color <- mylabcols[V(UKfaculty)$Group]

plot(UKfaculty)
```

We can also combine both options. This is usually what we do. 

```{r}
plot(UKfaculty,
     vertex.size = 9,
     # === vertex label properties
     vertex.label.family = "Helvetica",
     # === edge properties
     edge.width = 0.5,
     edge.arrow.size = 0.1)
```


### Coloring specifc edges

```{r}
E(UKfaculty)$color <- "gray"
E(UKfaculty)$color[E(UKfaculty)[47%--%28]] <- "red"

plot(UKfaculty, 
     # === vertex properties 
     vertex.size = 9,
     # === vertex label properties
     vertex.label.cex = 0.5,
     vertex.label.family = "Helvetica",
     # === edge properties
     edge.width = 0.5,
     edge.arrow.size = 0.1)

```

## Activity: 

Create your own plot of the karate data. Try changing different parameters to color different subsets of vertices and edges in a different way and write a short explanation of why you did what you did. 


```{r}
# Plot one 
# your code here
```

**Explaination: **

```{r}
# Plot two 
# your code here
```

**Explaination: **

```{r}
# Plot three
# your code here
```

**Explaination: **

```{r}
# Plot four 
# your code here
```


