---
title: "4 Activity: Descriptive Network Statistics (Metrics) : Centrality "
author: "Dr. Uma Ravat"
subtitle: '<br><br>Introduction to Social Network Analysis in R '
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


# 1. Setting up the network 

Load the `UKFaculty` Network

```{r setup_network}
# (remove # below and run this if they are not already installed )
# once installed put # back to avoid warnings/errors
#install.packages("igraph")
#install.packages("igraphdata")

library(igraphdata)
library(igraph)

## Some setup so plots are bigger with no margins
#default margins
def_marg <- c(5.1, 4.1, 4.1, 2.1)

#no margins
no_marg <- c(0, 0, 0, 0)

## load the data
data(UKfaculty)
```





### Activity 1: Descriptive statistics for the karate network : Centrality metrics

Explore centrality metrics for the `karate` network and interpret these metrics

_Be sure to load the network first_

_Add code chunks as necessary to run your code and answer the following_

`r bfcolor("Question:", "blue")`

```{r}
# Metrics
#Your code here
```

Your Interpretation of Metrics here: 



## Activity 2: Explore other centrality measures

[Other measures are in this link ](https://tidygraph.data-imaginist.com/reference/centrality.html)

Each group should pick three measures, discuss them and apply them to `UKFaculty` and `karate` network. 

